On the uniqueness and stability of dictionaries for sparse representation of noisy signals

### 2017-07-13 manuscripts/PNAS/chazthm_pnas.tex
newness
mathcal{H}

\documentclass[9pt,twocolumn]{pnas-new}
% Use the lineno option to display guide line numbers if required.
% Note that the use of elements such as single-column equations
% may affect the guide line number alignment.

%%% CHAZ TODO %%%

%%% CHRIS TODO %%
% steamroll

@@

% *** ALIGNMENT PACKAGES ***
\usepackage{array}
%\usepackage{cite}

@@

Here, we prove very generally that uniqueness and stability in sparse linear coding is an expected property of the model. 
More specifically, dictionaries that preserve sparse codes (i.e., satisfy a `spark condition') are identifiable from as few as \mbox{$N = m(k-1){m \choose k} + m$} noisy sparse linear combinations of their columns up to an error linear in the noise (Thm.~\ref{DeterministicUniquenessTheorem}). In fact, provided $n \geq \min(2k,m)$, in almost all cases the dictionary learning problem is well-posed (as per Hadamard \cite{Hadamard1902}) given enough data (Cor.~\ref{ProbabilisticCor}). Moreover, these guarantees hold without assuming the recovered matrix satisfies a spark condition, even when the number $m$ of dictionary elements is unknown. The explicit, algorithm-independent criteria we provide should be a useful theoretical tool.  %  in the theory of sparse representations.

@@

Before stating our main results, we first identify combinatorial criteria on the support sets of generating codes that imply stable sparse representations.  Letting $\{1, \ldots, m\}$ be denoted $[m]$, its power set $2^{[m]}$, and ${[m] \choose k}$ the set of subsets of $[m]$ of size $k$, we say a hypergraph $E \subseteq 2^{[m]}$ on vertices $[m]$ is \textit{$k$-uniform} when in fact $E \subseteq {[m] \choose k}$. The \emph{degree} $\deg(i)$ of a node $i \in [m]$ is the number of elements in $E$ which contain $i$, and we say $E$ is \emph{regular} when for some $r$ we have $\deg(i) = r$ for all $i$ (for given $r$, we say $E$ is \textit{$r$-regular}).

\begin{definition}\label{sip}
Given $E \subseteq 2^{[m]}$, the \textbf{star} $H(i)$ at $i$ is the set of $S \in E$ with $i \in S$. We say $E$ has the \textbf{singleton intersection property} (\textbf{SIP}) when $\cap H(i) = \{i\}$ for all $i \in [m]$. %We say $E$ has the SIP \textbf{of order $p$} when $\cap F(i) = \{i\}$ for every node $i$ for which $\text{deg(i)} \geq d_p$, where $d_1 \geq \ldots \geq d_m$ is the degree sequence of $E$.
\end{definition}

Next, we describe a quantitative version of the spark condition.  
The \emph{lower bound} $L$ of a matrix $\mathbf{M} \in \mathbb R^{n \times m}$ is the largest number $\alpha$ such that \mbox{$\|\mathbf{M}\mathbf{x}\|_2 \geq \alpha\|\mathbf{x}\|_2$} for all $\mathbf{x} \in \mathbb{R}^m$ \cite{Grcar10}. By compactness of the unit sphere, every injective linear map has a nonzero lower bound; hence, if $\mathbf{M}$ satisfies \eqref{SparkCondition}, then each submatrix formed from $2k$ of its columns or less has a nonzero lower bound. This motivates the following definition: 
%\begin{align*}
%L_k(\mathbf{M}) := \frac{1}{\sqrt{k}}\max \{ \alpha : \|\mathbf{M}\mathbf{x}\|_2 \geq \alpha\|\mathbf{x}\|_2 \text{ for all $k$-sparse } \mathbf{x}\}.
%\end{align*}
\begin{align*}
L_E(\mathbf{M}) := \inf \left\{ \frac{ \|\mathbf{M}_{S \cup S'}\mathbf{x}\|_2 }{ \sqrt{2k} \|\mathbf{x}\|_2} : \mathbf{x} \in \mathbb{R}^{|S\cup S'|}, \ \ S,S' \in E \right\}
%\frac{1}{\sqrt{k}}\max \{ \alpha : \|\mathbf{M}\mathbf{x}\|_2 \geq \alpha\|\mathbf{x}\|_2 \text{ for all $k$-sparse } \mathbf{x}\}.
\end{align*} 
%
and we write $L_{2k}:= L_E$ when $E = {[m] \choose k}$. Note that $ L = L_{\{[m]\}}$ whereas $L_{[m]} = L_2$. [TODO: Note when $L_2 > 0$ is implied by $L_E > 0$.]

Clearly, for any $\mathbf{M}$ satisfying \eqref{SparkCondition}, we have $L_{k'}(\mathbf{M}) > 0$ for  $k' \leq 2k$. We note that the quantity $1 - \sqrt{k} L_k(\mathbf{M})$ is also known in the compressive sensing literature as the (asymmetric) lower restricted isometry constant \cite{Blanchard2011}.

A vector $\mathbf{x}$ is said to be \emph{supported} on $S \subseteq [m]$ when $\mathbf{x} \in \text{\rm Span}( \{\mathbf{e}_j\}_{j\in S})$, where $\mathbf{e}_j$ are the standard basis in $\mathbb R^m$. For any index set $J$, denote by by $\mathbf{x}^J$ the subvector formed from the entries of $\mathbf{x}$ indexed by $J$, and similarly by $\mathbf{M}_J$ the submatrix formed by the columns of $\mathbf{M}$ indexed by $J$, where we set $\text{\rm Span}\{\mathbf{M}_\emptyset\} := \{\textbf{0}\}$. A set of $k$-sparse vectors is said to be in \emph{general linear position} when any $k$ of them are linearly independent. The following is a precise statement of our main result.

\begin{theorem}\label{DeterministicUniquenessTheorem}
Fix a real $n \times m$ matrix $\mathbf{A}$ with $L_E(\mathbf{A}) > 0$ for some $r$-regular $E \subseteq {[m] \choose k}$ with the SIP. If the sequence \mbox{$\mathbf{x}_1, \ldots, \mathbf{x}_N \in \mathbb{R}^m$} contains, for each $S \in E$, more than $(k-1){m \choose k}$ $k$-sparse vectors in general linear position supported on $S$, then there is a constant\footnote{We delay defining the constant $C_1$ until Section \ref{DUT} (\eqref{Cdef1}).} $C_1 > 0$ for which the following holds for all $\varepsilon < L_{2}(\mathbf{A}) / C_1$:

@@

Moreover, if in fact $L_{2k}(\mathbf{A}) > 0$ and $\varepsilon < L_{2k}(\mathbf{A}) / C_1$ then $L_{2k}(\mathbf{B}\mathbf{PD}) \geq L_{2k}(\mathbf{A}) - C_1 \varepsilon$ and:
\begin{align}\label{b-PDa}
\|\mathbf{x}^J_i - \mathbf{D}^{-1}\mathbf{P}^{\top}\mathbf{\bar x}^{\bar J}_i\|_1 &\leq  \left( \frac{ 1+C_1 \|\mathbf{x}^{J}_i\|_1 }{ L_{2k}(\mathbf{A}) -  C_1\varepsilon } \right) \varepsilon,\ \ \   \text{for $i \in [N]$}.
\end{align}
\end{theorem}
% You can recover matrices that don't satisfy the spark condition, but to uniquely recover k-sparse codes (since no limitation on their supports!) you need to have the spark condition. i guess...? Or can we use our knowledge of the particular permutation P to get rid of spark condition?

@@

%In fact, a more general result (stated clearly in the next section) can be gleaned from our method of proving Thm.~\ref{DeterministicUniquenessTheorem}. Briefly, in cases where $\mathbf{B}$ has $\bar m \neq m$ columns, or $E$ is not regular or only partially satisfying the SIP, a relation between $\bar m$ and the degree sequence of nodes in $E$ gives indices $J \subseteq [m]$ defining a submatrix $\mathbf{A}_J$ and subvectors $\mathbf{x}_i^J$ that are recoverable in the sense of \eqref{Cstable} and \eqref{b-PDa}. For example, if $E$ is $\ell$-regular with the SIP but $m \leq \bar m < m\ell/(\ell - 1)$ then we have nonzero $|J| = \bar m - \ell(\bar m - m)$. The implication here is that the smaller the difference $\bar m - m$, the more columns and code entries of the original $n \times m$ dictionary $\mathbf{A}$ and codes $\mathbf{x}_i$ contained (up to noise) in the appropriately scaled $n \times \bar m$ dictionary $\mathbf{B}$ and codes $\mathbf{\bar x}_i$. When $\bar m = m$, we recover Thm.~\ref{DeterministicUniquenessTheorem}.
Regarding the assumptions of Thm.~\ref{DeterministicUniquenessTheorem}, it is easy to verify that for every $k < m$, there is a regular $k$-uniform hypergraph that satisfies the SIP; for instance, take the consecutive intervals of length $k$ in some cyclic order on $[m]$.
In fact, even if $E$ is not regular or only partially satisfies the SIP, a relation between $\bar m$ and the degree sequence of nodes in $E$ may give the indices $J \subseteq [m]$. For sake of brevity, we delay to the next section a clear statement of this more general result.

@@

\begin{corollary}\label{DeterministicUniquenessCorollary}
Given $n, m$, $k < m$, and a regular hypergraph $E \subseteq {[m] \choose k}$ with the SIP, there are $N =  |E| \left[ (k-1){m \choose k} + 1  \right]$ vectors \mbox{$\mathbf{x}_1, \ldots, \mathbf{x}_N \in \mathbb{R}^m$} such that every matrix $\mathbf{A} \in \mathbb{R}^{n \times m}$ satisfying \eqref{SparkCondition} generates a dataset $Y = \{\mathbf{A}\mathbf{x}_1, \ldots, \mathbf{A}\mathbf{x}_N\}$ with a stable $k$-sparse representation in $\mathbb{R}^m$.
\end{corollary}

As already mentioned, there exist $k$-uniform regular hypergraphs $E$ with the SIP having cardinality $|E| = m$, implying the lower bound for sample size $N$ from the introduction. In many cases, however, the SIP can be achieved with far fewer supports; for example, when $k = \sqrt{m}$, take $E$ to be the $2k$ rows and columns formed by arranging $[m]$ in a square grid.

@@

\begin{corollary}\label{SLCopt}
If all of the assumptions of Thm.\ref{DeterministicUniquenessTheorem} hold with more than $(k-1)\left[ {\bar m \choose k} + |E|k{\bar m \choose k-1}\right]$ vectors $\mathbf{x}_i$ supported on each $S \in E$, then all solutions to Prob.~\ref{OptimizationProblem} necessarily satisfy the implications \eqref{Cstable} and \eqref{b-PDa} of Thm.~\ref{DeterministicUniquenessTheorem}.
\end{corollary}

@@

\begin{theorem}\label{robustPolythm}
There is a polynomial $g$ in the entries of $\mathbf{A}$ and $\mathbf{x}_i$ with the following property:  if $g$ evaluates to a nonzero number and more than \mbox{$(k-1){m \choose k}$} of the resulting $\mathbf{x}_i$ are supported on each $S \in E$ for some regular $E \subseteq {[m] \choose k}$ with the SIP, then $Y$ has a stable $k$-sparse representation in $\mathbb{R}^m$ (Def.~\ref{maindef}). In particular, all -- except for a Borel set of measure zero -- substitutions impart to $Y$ this property.
\end{theorem}

\begin{corollary}\label{ProbabilisticCor}
Fix $k < m$ and $n \geq \min(2k, m)$ and let the entries of $\mathbf{A} \in \mathbb{R}^{n \times m}$ and $k$-sparse $\mathbf{x}_1, \ldots, \mathbf{x}_N \in \mathbb{R}^m$ be drawn independently from probability measures absolutely continuous with respect to the standard Borel measure $\mu$. If more than $(k-1){m \choose k}$ of the vectors $\mathbf{x}_i$ are supported on each $S \in E$ for a regular $E \subseteq {[m] \choose k}$ with the SIP, then $Y$ has a stable $k$-sparse representation in $\mathbb{R}^m$ with probability one.
\end{corollary}

@@

As mentioned in the previous section, Thm.~\ref{DeterministicUniquenessTheorem} is a particular case of a more general result that forgoes the assumption that $E \subseteq {[m] \choose k}$ is regular and satisfies the SIP. If instead we require only that the stars $\cap H(i)$ intersect at singletons for all $i \leq q$ (assuming that the nodes of $E$ are labeled in some order of non-increasing degree), we have that $\bar m \geq k|E| / \deg(1)$ and, provided $\bar m < k|E| / (\deg(1) - 1)$, the non-empty submatrix $J$ is of size equal to the largest number $p$ satisfying:
\begin{align}\label{pcond}
\sum_{i=\ell}^{m} \deg(i) > (\bar m + 1 - \ell) (\deg(\ell) - 1) \ \ \text{for all } \ell \leq p \leq q.
\end{align}
Specifically, $J$ consists of the union of the set of all nodes of degree exceeding $\deg(p)$ and some subset of those nodes with degrees equal to $\deg(p)$. For the benefit of the reader, we do not prove this more general result below; it can be gleaned by examining how exactly Lemma \ref{NonEmptyLemma} is incorporated into the proof of Lem.~\ref{MainLemma}.
%As mentioned in the previous section, Thm.~\ref{DeterministicUniquenessTheorem} is a particular case of a more general result, which requires a looser set of constraints on the hypergraph $E$ and which applies to $n \times \bar m$ matrices $\mathbf{B}$ (and $\bar m$-dimensional codes $\mathbf{\bar x}_i$) with $\bar m \neq m$:

%Fix $\bar m$ and suppose the assumptions of Thm.~\ref{DeterministicUniquenessTheorem} hold, only now with the constraints on $E \subseteq {[m] \choose k}$ being just that $|\cap H(i)| = 1$ for all $i \leq q$ (assuming w.l.o.g. that the nodes of $E$ are labeled in some order of non-increasing degree), and with more than $(k-1){\bar m \choose k}$ vectors $\mathbf{x}_i$ supported in g.l.p. on each $S \in E$. Then we must have $\bar m \geq k|E| / \deg(1)$ and, provided $\bar m < k|E| / (\deg(1) - 1)$, the guarantee \eqref{Cstable} holds for a submatrix $\mathbf{A}_J$, where $J \subseteq[m]$ is nonempty and of a size equal to the largest number $p$ satisfying:
%\begin{align}\label{pcond}
%\sum_{i=\ell}^{m} \deg(i) > (\bar m + 1 - \ell) (\deg(\ell) - 1) \ \ \text{for all } \ell \leq p \leq q.
%\end{align}
%Specifically, $J$ consists of the union of the set of all nodes of degree exceeding $\deg(p)$ and some subset of those nodes with degrees equal to $\deg(p)$.

@@

The heart of the matter is therefore \eqref{Cstable}, which we now finally establish, first in the important special case $k = 1$.

@@

\begin{lemma}\label{MainLemma}
Suppose $\mathbf{A} \in \mathbb{R}^{n \times m}$ has $L_E(\mathbf{A}) > 0$ for some $r$-regular $E \subseteq {[m] \choose k}$ with the SIP. There exists $C_2 > 0$ for which the following holds for all $\varepsilon < L_2(\mathbf{A}) / C_2$:

If for some  $\mathbf{B} \in \mathbb{R}^{n \times \bar m}$ and map $\pi: E \mapsto {[\bar m] \choose k}$ we have:
\begin{align}\label{GapUpperBound}
d(\text{\rm Span}\{\mathbf{A}_{S}\}, \text{\rm Span}\{\mathbf{B}_{\pi(S)}\}) \leq \varepsilon, \ \  \text{for $S \in E$},
\end{align}
then $\bar m \geq m$, and provided $\bar m < mr /(r-1)$, there exists a permutation matrix $\mathbf{P}$ and invertible diagonal matrix $\mathbf{D}$ such that:
\begin{align}\label{MainLemmaBPD}
\|(\mathbf{A}_J)_j - \mathbf{B}_{\bar J} \mathbf{PD}_j\|_2 \leq C_2 \varepsilon, \ \  \text{for } j \in J,
\end{align}
for some non-empty $J, \bar J$ of size $\bar m - r(\bar m - m)$.
\end{lemma}

%\begin{lemma}\label{MainLemma}
%Fix $\mathbf{A} \in \mathbb{R}^{n \times m}$ with $L_E(\mathbf{A}) > 0$ for some regular $E \subseteq {[m] \choose k}$ with the SIP. There exists a constant $C_2 > 0$ for which the following holds for all $\varepsilon < L_2(\mathbf{A}) / C_2$:

%If a matrix $\mathbf{B} \in \mathbb{R}^{n \times m}$ and map $\pi: E \mapsto {m \choose k}$ satisfy:
%\begin{align}\label{GapUpperBound}
%d(\text{\rm Span}\{\mathbf{A}_{S}\}, \text{\rm Span}\{\mathbf{B}_{\pi(S)}\}) \leq \varepsilon, \ \ \text{for $S \in E$},
%\end{align}
%then there exist a permutation matrix $\mathbf{P}$ and invertible diagonal matrix $\mathbf{D}$ such that:
%\begin{align}\label{MainLemmaBPD}
%\|\mathbf{A}_j - \mathbf{BPD}_j\|_2 \leq C_2 \varepsilon, \ \  \text{for } j \in [m],
%\end{align}
%\end{lemma}

>>>>

\documentclass[9pt,twocolumn]{pnas-new}
% Use the lineno option to display guide line numbers if required.
% Note that the use of elements such as single-column equations may affect the guide line number alignment.

%%% TODO %%%
% Describe the connection between dictionary learning and union of subspace learning. Dictionaries are the top elements of the subspace intersection semi-lattice over reverse-inclusion. We identify subspaces generated by the elements of $A$ by showing that every $B$ must include these subspaces in its intersection semi-lattice. The dictionary elements are then inside these subspaces, so the lower the dimensionality of these subspaces the better.
% Shoutout to research on subspace arrangements? Not just "union of subspaces" model?
% Write Main Lemma or Lil lemma as a map between two intersection semi-lattices with a constraint (on the dimension of subspace? or on size of edge?)?
% Give example of when we might expect k-uniformity, i.e. x_i in glp on k-dimensional subspaces. Orbit under group action of a k-dimensional subspace?

@@

% *** ALIGNMENT PACKAGES ***
\usepackage{array}
\usepackage{bm}
%\usepackage{cite}

@@

Here, we prove very generally that uniqueness and stability in sparse linear coding is an expected property of the model. 
More specifically, dictionaries that preserve sparse codes (e.g., satisfy a `spark condition') are identifiable from as few as \mbox{$N = m(k-1){m \choose k} + m$} noisy sparse linear combinations of their columns up to an error linear in the noise (Thm.~\ref{DeterministicUniquenessTheorem}). In fact, provided $n \geq \min(2k,m)$, in almost all cases the dictionary learning problem is well-posed (as per Hadamard \cite{Hadamard1902}) given enough data (Cor.~\ref{ProbabilisticCor}). Moreover, these guarantees hold without assuming the recovered matrix satisfies a spark condition, even when the number $m$ of dictionary elements is unknown. The explicit, algorithm-independent criteria we provide should be a useful theoretical tool.  %  in the theory of sparse representations.

@@

Before stating our main result, we identify combinatorial criteria on the support sets of the generating codes $\mathbf{x}_i$ that imply stable sparse representations. Letting $\{1, \ldots, m\}$ be denoted $[m]$, its power set $2^{[m]}$, and ${[m] \choose k}$ the set of subsets of $[m]$ of size $k$, we say a hypergraph $\mathcal{H} \subseteq 2^{[m]}$ on vertices $[m]$ is \textit{$k$-uniform} when in fact $\mathcal{H} \subseteq {[m] \choose k}$. The \emph{degree} $\deg(i)$ of a node $i \in [m]$ is the number of sets in $\mathcal{H}$ that contain $i$, and we say $\mathcal{H}$ is \emph{regular} when for some $r$ we have $\deg(i) = r$ for all $i$ (for given $r$, we say $\mathcal{H}$ is \textit{$r$-regular}).

\begin{definition}\label{sip}
Given $\mathcal{H} \subseteq 2^{[m]}$, the \textbf{star} $\mathcal{G}(i)$ at $i$ is the set of $S \in \mathcal{H}$ with $i \in S$. We say $\mathcal{H}$ has the \textbf{singleton intersection property} (\textbf{SIP}) when $\cap \mathcal{G}(i) = \{i\}$ for all $i \in [m]$.
\end{definition}

Next, we describe a quantitative version of the spark condition \eqref{SparkCondition}. The \emph{lower bound} $L$ of a matrix $\mathbf{M} \in \mathbb R^{n \times m}$ is the largest number $\alpha$ such that \mbox{$\|\mathbf{M}\mathbf{x}\|_2 \geq \alpha\|\mathbf{x}\|_2$} for all $\mathbf{x} \in \mathbb{R}^m$ \cite{Grcar10}. By compactness of the unit sphere, every injective linear map has a nonzero lower bound; hence, if $\mathbf{M}$ satisfies \eqref{SparkCondition}, then each submatrix formed from $2k$ of its columns or less has a nonzero lower bound.

We generalize this domain-restricted lower bound to a \emph{union of subspaces} model \cite{lu2008theory, vidal2005generalized} derived from the hypergraph $\mathcal{H}$. Let $\mathbf{M}_J$ be the submatrix formed by the columns of $\mathbf{M}$ indexed by $J$ and let $\bm{\mathcal{M}}_J := \text{span}\{\mathbf{M}_J\}$, with $\bm{\mathcal{M}}_\emptyset := \mathbf{0}$. We define: 
\begin{align*}
L_\mathcal{H}(\mathbf{M}) := \inf \left\{ \frac{ \|\mathbf{M}(\mathbf{x}_1-\mathbf{x}_2)\|_2 }{ \sqrt{2k} \|\mathbf{x}_1-\mathbf{x}_2\|_2} : \mathbf{x}_1, \mathbf{x}_2 \in \cup_{S \in \mathcal{H}} \bm{\mathcal{M}}_S \right\},
%\frac{1}{\sqrt{k}}\max \{ \alpha : \|\mathbf{M}\mathbf{x}\|_2 \geq \alpha\|\mathbf{x}\|_2 \text{ for all $k$-sparse } \mathbf{x}\}.
\end{align*} 
%
%This motivates the following definition: 
%\begin{align*}
%L_k(\mathbf{M}) := \frac{1}{\sqrt{k}}\max \{ \alpha : \|\mathbf{M}\mathbf{x}\|_2 \geq \alpha\|\mathbf{x}\|_2 \text{ for all $k$-sparse } \mathbf{x}\}.
%\end{align*}
%\begin{align*}
%L_\mathcal{H}(\mathbf{M}) := \inf \left\{ \frac{ \|\mathbf{M}_{S \cup S'}\mathbf{x}\|_2 }{ \sqrt{2k} \|\mathbf{x}\|_2} : \mathbf{x} \in \mathbb{R}^{|S\cup S'|}, \ \ S,S' \in \mathcal{H} \right\}
%\frac{1}{\sqrt{k}}\max \{ \alpha : \|\mathbf{M}\mathbf{x}\|_2 \geq \alpha\|\mathbf{x}\|_2 \text{ for all $k$-sparse } \mathbf{x}\}.
%\end{align*} 
%
where we write $L_{2k}:= L_\mathcal{H}$ when $\mathcal{H} = {[m] \choose k}$. Beware that $ L = L_{\{[m]\}}$ whereas $L_{[m]} = L_2$. [TODO: Note when $L_2 > 0$ is implied by $L_\mathcal{H} > 0$.] We note that the quantity $1 - \sqrt{k} L_k(\mathbf{M})$ is also known in the compressive sensing literature as the (asymmetric) lower restricted isometry constant \cite{Blanchard2011}. Clearly, for any $\mathbf{M}$ satisfying \eqref{SparkCondition}, we have $L_{k'}(\mathbf{M}) > 0$ for  $k' \leq 2k$.

A vector $\mathbf{x}$ is said to be \emph{supported} on $S \subseteq [m]$ when $\mathbf{x} \in \text{\rm span}\{\mathbf{e}_j: j\in S\}$, where $\mathbf{e}_j$ are the standard basis in $\mathbb R^m$. Denote by $\mathbf{x}^J$ the subvector formed from the entries of $\mathbf{x}$ indexed by $J$. A set of $k$-sparse vectors is said to be in \emph{general linear position} when any $k$ of them are linearly independent. The following is a precise statement of our main result.

\begin{theorem}\label{DeterministicUniquenessTheorem}
Fix a real $n \times m$ matrix $\mathbf{A}$ with $L_\mathcal{H}(\mathbf{A}) > 0$ for some $r$-regular $\mathcal{H} \subseteq {[m] \choose k}$ with the SIP. If the sequence \mbox{$\mathbf{x}_1, \ldots, \mathbf{x}_N \in \mathbb{R}^m$} contains, for each $S \in \mathcal{H}$, more than $(k-1){m \choose k}$ $k$-sparse vectors in general linear position supported on $S$, then there is a constant\footnote{We delay defining the constant $C_1$ until Section \ref{DUT} (\eqref{Cdef1}).} $C_1 > 0$ for which the following holds for all $\varepsilon < L_{2}(\mathbf{A}) / C_1$:

@@

Moreover, if in fact $L_{2k}(\mathbf{A}) > 0$ and $\varepsilon < L_{2k}(\mathbf{A}) / C_1$ then $L_{2k}(\mathbf{B}\mathbf{PD}) \geq L_{2k}(\mathbf{A}) - C_1 \varepsilon$ and:
\begin{align}\label{b-PDa}
\|\mathbf{x}^J_i - \mathbf{D}^{-1}\mathbf{P}^{\top}\mathbf{\bar x}^{\bar J}_i\|_1 &\leq  \left( \frac{ 1+C_1 \|\mathbf{x}^{J}_i\|_1 }{ L_{2k}(\mathbf{A}) -  C_1\varepsilon } \right) \varepsilon \ \  \text{for $i \in [N]$}.
\end{align}
\end{theorem}
% You can recover matrices that don't satisfy the spark condition, but to uniquely recover k-sparse codes (since no limitation on their supports!) you need to have the spark condition. i guess...? Or can we use our knowledge of the particular permutation P to get rid of spark condition?

@@

%In fact, a more general result (stated clearly in the next section) can be gleaned from our method of proving Thm.~\ref{DeterministicUniquenessTheorem}. Briefly, in cases where $\mathbf{B}$ has $\bar m \neq m$ columns, or $\mathcal{H}$ is not regular or only partially satisfying the SIP, a relation between $\bar m$ and the degree sequence of nodes in $\mathcal{H}$ gives indices $J \subseteq [m]$ defining a submatrix $\mathbf{A}_J$ and subvectors $\mathbf{x}_i^J$ that are recoverable in the sense of \eqref{Cstable} and \eqref{b-PDa}. For example, if $\mathcal{H}$ is $\ell$-regular with the SIP but $m \leq \bar m < m\ell/(\ell - 1)$ then we have nonzero $|J| = \bar m - \ell(\bar m - m)$. The implication here is that the smaller the difference $\bar m - m$, the more columns and code entries of the original $n \times m$ dictionary $\mathbf{A}$ and codes $\mathbf{x}_i$ contained (up to noise) in the appropriately scaled $n \times \bar m$ dictionary $\mathbf{B}$ and codes $\mathbf{\bar x}_i$. When $\bar m = m$, we recover Thm.~\ref{DeterministicUniquenessTheorem}.
Regarding the assumptions of Thm.~\ref{DeterministicUniquenessTheorem}, it is easy to verify that for every $k < m$, there is a regular $k$-uniform hypergraph that satisfies the SIP; for instance, take the consecutive intervals of length $k$ in some cyclic order on $[m]$.
In fact, even if $\mathcal{H}$ is not regular or only partially satisfies the SIP, a relation between $\bar m$ and the degree sequence of nodes in $\mathcal{H}$ may give the indices $J \subseteq [m]$. For sake of brevity, we delay to the next section a clear statement of this more general result.

@@

\begin{corollary}\label{DeterministicUniquenessCorollary}
Given $n, m$, $k < m$, and a regular hypergraph $\mathcal{H} \subseteq {[m] \choose k}$ with the SIP, there are $N =  |\mathcal{H}| \left[ (k-1){m \choose k} + 1  \right]$ vectors \mbox{$\mathbf{x}_1, \ldots, \mathbf{x}_N \in \mathbb{R}^m$} such that every matrix $\mathbf{A} \in \mathbb{R}^{n \times m}$ satisfying \eqref{SparkCondition} generates a dataset $Y = \{\mathbf{A}\mathbf{x}_1, \ldots, \mathbf{A}\mathbf{x}_N\}$ with a stable $k$-sparse representation in $\mathbb{R}^m$.
\end{corollary}

As already mentioned, there exist $k$-uniform regular hypergraphs $\mathcal{H}$ with the SIP having cardinality $|\mathcal{H}| = m$, implying the lower bound for sample size $N$ from the introduction. In many cases, however, the SIP can be achieved with far fewer supports; for example, when $k = \sqrt{m}$, take $\mathcal{H}$ to be the $2k$ rows and columns formed by arranging $[m]$ in a square grid.

@@

\begin{corollary}\label{SLCopt}
If all of the assumptions of Thm.\ref{DeterministicUniquenessTheorem} hold with more than $(k-1)\left[ {\bar m \choose k} + |\mathcal{H}|k{\bar m \choose k-1}\right]$ vectors $\mathbf{x}_i$ supported on each $S \in \mathcal{H}$, then all solutions to Prob.~\ref{OptimizationProblem} necessarily satisfy the implications \eqref{Cstable} and \eqref{b-PDa} of Thm.~\ref{DeterministicUniquenessTheorem}.
\end{corollary}

@@

\begin{theorem}\label{robustPolythm}
There is a polynomial $g$ in the entries of $\mathbf{A}$ and $\mathbf{x}_i$ with the following property:  if $g$ evaluates to a nonzero number and more than \mbox{$(k-1){m \choose k}$} of the resulting $\mathbf{x}_i$ are supported on each $S \in \mathcal{H}$ for some regular $\mathcal{H} \subseteq {[m] \choose k}$ with the SIP, then $Y$ has a stable $k$-sparse representation in $\mathbb{R}^m$ (Def.~\ref{maindef}). In particular, all -- except for a Borel set of measure zero -- substitutions impart to $Y$ this property.
\end{theorem}

\begin{corollary}\label{ProbabilisticCor}
Fix $k < m$ and $n \geq \min(2k, m)$ and let the entries of $\mathbf{A} \in \mathbb{R}^{n \times m}$ and $k$-sparse $\mathbf{x}_1, \ldots, \mathbf{x}_N \in \mathbb{R}^m$ be drawn independently from probability measures absolutely continuous with respect to the standard Borel measure $\mu$. If more than $(k-1){m \choose k}$ of the vectors $\mathbf{x}_i$ are supported on each $S \in \mathcal{H}$ for a regular $\mathcal{H} \subseteq {[m] \choose k}$ with the SIP, then $Y$ has a stable $k$-sparse representation in $\mathbb{R}^m$ with probability one.
\end{corollary}

@@

As mentioned in the previous section, Thm.~\ref{DeterministicUniquenessTheorem} is a particular case of a more general result that forgoes the assumption that $\mathcal{H} \subseteq {[m] \choose k}$ is regular and satisfies the SIP. If instead we require only that the stars $\cap H(i)$ intersect at singletons for all $i \leq q$ (assuming that the nodes of $\mathcal{H}$ are labeled in some order of non-increasing degree), we have that $\bar m \geq k|\mathcal{H}| / \deg(1)$ and, provided $\bar m < k|\mathcal{H}| / (\deg(1) - 1)$, the non-empty submatrix $J$ is of size equal to the largest number $p$ satisfying:
\begin{align}\label{pcond}
\sum_{i=\ell}^{m} \deg(i) > (\bar m + 1 - \ell) (\deg(\ell) - 1) \ \ \text{for all } \ell \leq p \leq q.
\end{align}
Specifically, $J$ consists of the union of the set of all nodes of degree exceeding $\deg(p)$ and some subset of those nodes with degrees equal to $\deg(p)$. For the benefit of the reader, we do not prove this more general result below; it can be gleaned by examining how exactly Lemma \ref{NonEmptyLemma} is incorporated into the proof of Lem.~\ref{MainLemma}.
%As mentioned in the previous section, Thm.~\ref{DeterministicUniquenessTheorem} is a particular case of a more general result, which requires a looser set of constraints on the hypergraph $\mathcal{H}$ and which applies to $n \times \bar m$ matrices $\mathbf{B}$ (and $\bar m$-dimensional codes $\mathbf{\bar x}_i$) with $\bar m \neq m$:

%Fix $\bar m$ and suppose the assumptions of Thm.~\ref{DeterministicUniquenessTheorem} hold, only now with the constraints on $\mathcal{H} \subseteq {[m] \choose k}$ being just that $|\cap H(i)| = 1$ for all $i \leq q$ (assuming w.l.o.g. that the nodes of $\mathcal{H}$ are labeled in some order of non-increasing degree), and with more than $(k-1){\bar m \choose k}$ vectors $\mathbf{x}_i$ supported in g.l.p. on each $S \in \mathcal{H}$. Then we must have $\bar m \geq k|\mathcal{H}| / \deg(1)$ and, provided $\bar m < k|\mathcal{H}| / (\deg(1) - 1)$, the guarantee \eqref{Cstable} holds for a submatrix $\mathbf{A}_J$, where $J \subseteq[m]$ is nonempty and of a size equal to the largest number $p$ satisfying:
%\begin{align}\label{pcond}
%\sum_{i=\ell}^{m} \deg(i) > (\bar m + 1 - \ell) (\deg(\ell) - 1) \ \ \text{for all } \ell \leq p \leq q.
%\end{align}
%Specifically, $J$ consists of the union of the set of all nodes of degree exceeding $\deg(p)$ and some subset of those nodes with degrees equal to $\deg(p)$.

@@

The heart of the matter is therefore \eqref{Cstable}, which we now finally establish, first in the important special case $k = 1$

@@

\begin{lemma}\label{MainLemma}
Suppose $\mathbf{A} \in \mathbb{R}^{n \times m}$ has $L_\mathcal{H}(\mathbf{A}) > 0$ for some $r$-regular $\mathcal{H} \subseteq {[m] \choose k}$ with the SIP. There exists $C_2 > 0$ for which the following holds for all $\varepsilon < L_2(\mathbf{A}) / C_2$:

If for some  $\mathbf{B} \in \mathbb{R}^{n \times \bar m}$ and map $\pi: E \mapsto {[\bar m] \choose k}$ we have:
\begin{align}\label{GapUpperBound}
d(\bm{\mathcal{A}}_S, \bm{\mathcal{B}}_{\pi(S)}) \leq \varepsilon, \ \  \text{for $S \in \mathcal{H}$},
\end{align}
then $\bar m \geq m$, and provided $\bar m < mr /(r-1)$, there exists a permutation matrix $\mathbf{P}$ and invertible diagonal matrix $\mathbf{D}$ such that:
\begin{align}\label{MainLemmaBPD}
\|(\mathbf{A}_J)_j - \mathbf{B}_{\bar J} \mathbf{PD}_j\|_2 \leq C_2 \varepsilon, \ \  \text{for } j \in J,
\end{align}
for some non-empty $J, \bar J$ of size $\bar m - r(\bar m - m)$.
\end{lemma}

%\begin{lemma}\label{MainLemma}
%Fix $\mathbf{A} \in \mathbb{R}^{n \times m}$ with $L_\mathcal{H}(\mathbf{A}) > 0$ for some regular $\mathcal{H} \subseteq {[m] \choose k}$ with the SIP. There exists a constant $C_2 > 0$ for which the following holds for all $\varepsilon < L_2(\mathbf{A}) / C_2$:

%If a matrix $\mathbf{B} \in \mathbb{R}^{n \times m}$ and map $\pi: E \mapsto {m \choose k}$ satisfy:
%\begin{align}\label{GapUpperBound}
%d(\text{\rm span}\{\mathbf{A}_{S}\}, \bm{\mathcal{B}}_{\pi(S)}) \leq \varepsilon, \ \ \text{for $S \in \mathcal{H}$},
%\end{align}
%then there exist a permutation matrix $\mathbf{P}$ and invertible diagonal matrix $\mathbf{D}$ such that:
%\begin{align}\label{MainLemmaBPD}
%\|\mathbf{A}_j - \mathbf{BPD}_j\|_2 \leq C_2 \varepsilon, \ \  \text{for } j \in [m],
%\end{align}
%\end{lemma}
